---
title: "Project 2.2"
author: "Truc Le"
date: "3/1/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(summarytools)
library(dplyr)
library(coin)
library(ggpubr)
setwd("/Users/trucle/Desktop")
covidB <- read.csv("CovidB.csv")
covidB$Age_Group <-as.factor(covidB$Age_Group)

minor <- subset(covidB, Age_Group == "0-17 years")
Y_adult <- subset(covidB, Age_Group == "45-54 years")
elderly <- subset(covidB, Age_Group == "75-84 years")

data <- rbind(minor, Y_adult, elderly)

 duptimes <- c( 6,1,7,9,13,6,11,6,18,11,8,8,9,4,6,7,13,11,14,11,15,8,0,2,375,145,2252,926,1179,527,757,381,1305,680,1123,590,625,343,667,336,1214,661,2475,1276,2332,1254,491,253,1105,795,9791,7833,5148,4617,2317,2158,4568,3453,4460,3577,2916,2304,3923,3076,8749,6427,15263,11571,13872,10493,3014,2158)
 
 idx <- rep(1:nrow(data), duptimes)
 
 dupdf <- data[idx,]
 
 data_new <-dupdf[3:4]
```

#1 Introduction 



#2 Summary Analysis

```{r, echo=FALSE}
plot(Death~Age_Group, data=data)

par(mfrow=c(2,2))

a<-ggplot(data, aes(x = Age_Group, y = Death))+
  theme_bw()+
  geom_bar(stat = "identity")+
  ggtitle("Barplot for Death Counts by Age Groups")+
  xlab("Age Group")+
  ylab("Death Counts")

b<-ggplot(data, aes(x = Sex, y = Death))+
  theme_bw()+
  geom_bar(stat = "identity")+
  ggtitle("Barplot for Death Counts by Gender")+
  xlab("Gender")+
  ylab("Death Counts")

c<-ggplot(data, aes(x = Age_Group, y = Death))+
  theme_bw()+
  geom_boxplot()+
  ggtitle("Boxplot for Death Counts by Age Groups")+
  xlab("Age Group")+
  ylab("Death Counts")

d<-ggplot(data, aes(x = Sex, y = Death))+
  theme_bw()+
  geom_boxplot()+
  ggtitle("Boxplot for Death Counts by Gender")+
  xlab("Gender")+
  ylab("Death Counts")

text <- paste("Figure 1: From the bar graphs and boxplots, it can be observe that the death counts are the highest in the age group of 75-85 years and in the male population", sep = " ")
text.p <- ggparagraph(text = text, face = "italic", size = 11, color = "black")

ggarrange(a, b, c, d, text.p, 
          labels = c("A", "B", "C","D"),
          ncol = 2, nrow = 3)

# with(data, tapply(Death, Age_Group, mean))
# with(data, tapply(Death, Age_Group, median))
# with(data, tapply(Death, Age_Group, sd))
# with(data, tapply(Death, Age_Group, min))
# with(data, tapply(Death, Age_Group, max))
# with(data, tapply(Death, Age_Group, sum))
# 
with(data, tapply(Death, Sex, mean))
with(data, tapply(Death, Sex, median))
with(data, tapply(Death, Sex, sd))
with(data, tapply(Death, Sex, min))
with(data, tapply(Death, Sex, max))
with(data, tapply(Death, Sex, sum))
```

###Statistical Summary Table 

| Death Counts by: | 0-17 years | 45-54 years | 75-84 years | Female   | Male     |
|------------------|------------|-------------|-------------|----------|----------|
| Mean             | 8.50000    | 923.62500   | 5564.66667  | 1830.056 | 2501.139 |
| Median           | 8.0        | 673.5       | 4191.5      | 558.5    | 1114.0   |
| Std Dev          | 4.423259   | 653.541377  | 4058.287448 | 2941.454 | 3793.517 |
| Min              | 0          | 145         | 759         | 1        | 0        |
| Max              | 18         | 2475        | 15263       | 11571    | 15263    |
| total            | 204        | 22167       | 133552      | 65882    | 90041    |
Table 1: From this sample, it is observed that the age group of 75-84 years have the highest mortality rate from Covid-19


#3 Analysis

```{r, echo=FALSE}
data_table<-xtabs(Death ~ Sex + Age_Group, data = data,  drop.unused.levels = TRUE)
ni. = rowSums(data_table)
n.j = colSums(data_table)

#class(data_table)
```

Hypothesis:

$H_O:$ Age Group and Genders are indepedent variables regarding Death involving Covid-19.

$H_A:$ Age Group and Genders are not indepedent variables regarding Death involving Covid-19.


###Contingency Table
|        | Age Group  |             |             |             
|--------|------------|-------------|-------------|
| Sex    | 0-17 years | 45-54 years | 75-84 years |
| Female | 84         | 7372        | 58426       |
| Male   | 120        | 14795       | 75126       |
Table 2: It is observed that there is a higher mortality rate in the male population for each Age Group.



```{r, echo=FALSE}
the.test = chisq.test(data_table,correct = FALSE)
eij = the.test$expected
chi.sq.obs = as.numeric(the.test$statistic)
```

```{r, echo=FALSE}
#Permutation P-value
R = 4000
r.perms = sapply(1:R,function(i){
  perm.data = data
  perm.data$Age_Group = sample(perm.data$Age_Group,nrow(perm.data),replace = FALSE)
  chi.sq.i = chisq.test(xtabs(Death~Sex+ Age_Group, data=perm.data, drop.unused.levels = TRUE),correct = FALSE)$stat
  return(chi.sq.i)
})
perm.pval = mean(chi.sq.i >= chi.sq.obs)
```


P( 0-17 years | Female) - P( 0-17 years | Male)
P(18-29 years | Female) - P(18-29 years | Male)
P(45-54 years | Female) - P(45-54 years | Male)
P(75-84 years | Female) - P(75-84 years | Male)

```{r, echo=FALSE}
#Pairwise comparisons

n = sum(data_table)
ni. = rowSums(data_table)
n.j = colSums(data_table)
all.pjG1 = data_table[1,]/ni.[1] #all conditional probabilites for row 1
all.pjG2= data_table[2,]/ni.[2] #all conditional probabilites for row 2
all.pbar = n.j/n #all probabilities regardless of group
all.Zij = c(all.pjG1 - all.pjG2)/sqrt(all.pbar*(1-all.pbar)*(1/ni.[1] + 1/ni.[2])) #The z-test-statistics

r.perms.cutoff = sapply(1:R,function(i){
  perm.data = data
  perm.data$Sex = sample(perm.data$Sex,nrow(perm.data),replace = FALSE)
  row.sum = rowSums(xtabs(Death~Sex+ Age_Group, data=perm.data, drop.unused.levels = TRUE))
  col.sum = colSums(xtabs(Death~Sex+ Age_Group, data=perm.data, drop.unused.levels = TRUE))
  all.pji = xtabs(Death~Sex+ Age_Group, data=perm.data, drop.unused.levels = TRUE)[1,]/row.sum[1]
  all.pji.= xtabs(Death~Sex+ Age_Group, data=perm.data, drop.unused.levels = TRUE)[2,]/row.sum[2]
  all.pbar = col.sum/sum(row.sum)
  all.Zij = c(all.pji - all.pji.)/sqrt(all.pbar*(1-all.pbar)*(1/row.sum[1] + 1/row.sum[2]))
  Q.r = max(abs(all.Zij))
  return(Q.r)
})
alpha = 0.05
cutoff.q = as.numeric(quantile(21.55774,(1-alpha)))



all.Zij = matrix(all.Zij,nrow=  1)
colnames(all.Zij) = c("0-17 years","45-54 years", "75-84 years")
rownames(all.Zij) = c("Female vs. Male")
all.Zij
```